---
title: "Analysis of annuals in Australia"
format: html
editor: visual
---

# Background & aims

## Aims

1.  Replicate and extend results of Poppenwimer et al 2023 (https://doi.org/10.1101/2022.09.13.507878) showing annuals are more common in dry, unpredictable envs in Australia

# Datasets

## AusTraits

Load Austraits growth form (using only public released data)

```{r}
#install.packages("tidyverse")
library(tidyverse)

#remotes::install_github("traitecoevo/austraits")
library(austraits) 

# load austraits
austraits <- load_austraits(path="data/austraits", version = get_version_latest())

# filter to the life history data and remove problematic datasets
lh_taxa <- austraits$traits %>% filter(trait_name == "life_history", 
                                       !dataset_id %in% c("Sjostrom_2006", "ABRS_1981", "WAH_1998", "WAH_2016", "NHNSW_2014", "NHNSW_2014_2"))

# I can provide code for a quick-and-dirty way to get just one consensus value per species if needed

```

## GBIF

Load cleaned GBIF/ALA dataset as tibble

```{r}

```

## Spatial grid

Create/load a grid of cells in Australia, that we will use for analysis

```{r}

```

How to identify which cell each point is located in

```{r}

```

## Coavariates

Download climate, soil and other covariates

```{r}

```
